Real-Time Activity Classification by Match Filtering using Body-Worn Accelerometers

### 2016-07-03 notes/paper/paper.tex
Spell checking to the paper.

\documentclass[journal]{IEEEtran}
\usepackage{blindtext}
\usepackage{graphicx}
\usepackage{subfig}
\usepackage[english]{babel}
\usepackage{pinyin}
\usepackage{natbib}
\usepackage{notoccite}
\usepackage{amssymb}
\usepackage{textcomp}
\usepackage{floatrow}
\usepackage{amsmath}
%
\begin{document}
%
% paper title
\title{Real-Time Activity Classification by Match Filtering using Body-Worn Accelerometers}
%
\author{Craig~Euler, Bryan~Juarez, Melissa~Flores, C.T.~Lin}
%
\maketitle
%
\begin{abstract}
Having information of the user\textquotesingle s activity can provide great use in modern-day devices such as ones that track and monitor a user\textquotesingle s activity and fitness.
In this paper, we demonstrate activity classification performance of the matched-filtering method with data obtained from a three-axis accelerometer worn by the user.
We also show the real-time processing capability of our algorithms on the MSP432P401R microcontroller.
Dimensionality reduction with principal component analysis (PCA) \cite{bishop_2006} is a data compression technique we use to improve our processing throughput which, inherently, has the added benifit of making our data invarient to sensor orientation when applied.
Data decimation in time is an additional throughput enhancement that we apply early to our data.
We make use of an instance-based learning algorithm to train the device to learn the individual\textquotesingle s motion patterns and store that information as activity templates for use in our matched-filter.
\end{abstract}
%
\section{Introduction}
Various methods for activity classification using body-worn sensors exist.
Custom Decision Tree, Automatically Generated Decision Tree, and Artificial Neural Networks \cite{parkka_ermes_korpipaa_mantyjarvi_peltola_korhonen_2006} have been explored as well as other frequency based classification methods \cite{sharma_purwar_lee_lee_chung_2008}.
In this paper, we make use of the matched-filter method to identify an individual\textquotesingle s activity for the purpose of real-time processing.
The matched-filter method with use of thresholding and comparing the correlations of various filters is a simple method to use and has been studied in the past \cite{giannakis_tsatsanis_1990}.
The methods used for our algorithms are designed for computational efficiency to accommodate the resource limitations of a microcontroller (we use the MSP432 microcontroller).

@@

The real-time activity classification module is intended to be performed on a microcontroler.
This module reads in the references stored on the unit to be used as templates that the matched-filter will use for determining the individual\textquotesingle s activity.
Both of these modules use the same principles and hence the same algorithm kernels.
%
\section{The Methodology}
%
\subsection{Activity Classification Method}
The matched-filter method is commonly used for extracting a signal out of noisy data.
We use this method for identifying a user\textquotesingle s activity from data generated by body-worn accelerometers.
Because the matched-filter method requires use of an existing reference template to identify the desired signal, we must develop our template from a training set.
We will discuss how we develop our set of templates for use in our matched-filter in section \ref{sec:training}.

@@

\begin{figure}[!ht]
   \centering
   \subfloat[\label{fig:multiple_comparison}]{\includegraphics[width=0.45\textwidth]{more_time_plots.png}}
   \quad
   \subfloat[\label{fig:timing_refs}]{\includegraphics[width=0.45\textwidth]{one_and_two_references.png}}
   \caption{Timing results with varying sampling frequencies, time-window lengths, and number of references used for the matched-filter on the MSP432 microcontroller processor. The time window used for (b) is 4.0 seconds with timings measured at sampling frequencies of $16, 32, 62, 64, 75, 100,$ and $128$ Hz.}
   \label{fig:timing}
\end{figure}
%
Because the accelerometer was not connected to the microcontroller at the time of writing this paper, randomized data was used for the purpose of analyzing the throughput performance on the device while varying the parameters which accurately models the processing load we would expect when we do have the accelerometer connected to the microcontroller.
Figure \ref{fig:multiple_comparison} shows the execution time with varying time-window lengths and sampling frequencies while maintaining two references used in the matched-filter process and operating at decimation $2$.
Because the FFT dominates the processing load, and was optimized for even-powered factors, one expects best throughput at radix 2 valued time-window lengths and sampling frequencies followed by other even-valued, non-radix 2, time-lengths and sampling frequencies. Both $62$Hz and $50$Hz have only one factor of $2$ hence the expectation that a sampling frequency of $62$Hz processing time is longer than that the $50$Hz.
Figure \ref{fig:timing_refs} shows the execution time over a $4.0$ second time window at decimation $8$ while varying the sampling frequency and number of references. Because the pre-processing execution time does not change with the number of references, it's observed that matched-filtering accounts for roughly $80$\% of the processing while pre-processing accounts for the other $20$\% when processing for a single reference.
%
\section{Conclusion}
We have demonstrated excellent activity classification capability of the matched-filter method using data obtained from a Kinetisense tri-axis accelerometer sensor.
We have shown that real-time processing can be achieved with timing results obtained from the MSP432 microcontroller.
Timing results have demonstrated that radix 2 time-window lengths and frequency values result in faster processing times as expected given that the FFT dominates the procesing time.
We increased our throughput with the use of data-compression techniques, namely decimation and dimensionality reduction, which reduced our computational requirements.
The references used in the matched-filter are obtained from the training module where a training set is provided by the user through some specified activity and the best representative motion signature of the individual is extracted through the use of an instance-based learning algorithm.

>>>>

\documentclass[journal]{IEEEtran}
\usepackage{blindtext}
\usepackage{graphicx}
\usepackage{subfig}
\usepackage[english]{babel}
\usepackage{pinyin}
\usepackage{natbib}
\usepackage{notoccite}
\usepackage{amssymb}
\usepackage{textcomp}
\usepackage{floatrow}
\usepackage{amsmath}
%
\begin{document}
%
% paper title
\title{Real-Time Activity Classification by Match Filtering using Body-Worn Accelerometers}
%
\author{Craig~Euler, Bryan~Juarez, Melissa~Flores, C.T.~Lin}
%
\maketitle
%
\begin{abstract}
Having information of the user\textquotesingle s activity can provide great use in modern-day devices such as ones that track and monitor a user\textquotesingle s activity and fitness.
In this paper, we demonstrate activity classification performance of the matched-filtering method with data obtained from a three-axis accelerometer worn by the user.
We also show the real-time processing capability of our algorithms on the MSP432P401R microcontroller.
Dimensionality reduction with principal component analysis (PCA) \cite{bishop_2006} is a data compression technique we use to improve our processing throughput which, inherently, has the added benefit of making our data invariant to sensor orientation when applied.
Data decimation in time is an additional throughput enhancement that we apply early to our data.
We make use of an instance-based learning algorithm to train the device to learn the individual\textquotesingle s motion patterns and store that information as activity templates for use in our matched-filter.
\end{abstract}
%
\section{Introduction}
Various methods for activity classification using body-worn sensors exist.
Custom Decision Tree, Automatically Generated Decision Tree, and Artificial Neural Networks \cite{parkka_ermes_korpipaa_mantyjarvi_peltola_korhonen_2006} have been explored as well as other frequency based classification methods \cite{sharma_purwar_lee_lee_chung_2008}.
In this paper, we make use of the matched-filter method to identify an individual\textquotesingle s activity for the purpose of real-time processing.
The matched-filter method with use of thresholding and comparing the correlations of various filters is a simple method to use and has been studied in the past \cite{giannakis_tsatsanis_1990}.
The methods used for our algorithms are designed for computational efficiency to accommodate the resource limitations of a microcontroller (we use the MSP432 microcontroller).

@@

The real-time activity classification module is intended to be performed on a microcontroller.
This module reads in the references stored on the unit to be used as templates that the matched-filter will use for determining the individual\textquotesingle s activity.
Both of these modules use the same principles and hence the same algorithm kernels.
%
\section{The Methodology}
%
\subsection{Activity Classification Method}
The matched-filter method is commonly used for extracting a signal out of noisy data.
We use this method for identifying a user\textquotesingle s activity from data generated by body-worn accelerometers.
Because the matched-filter method requires use of an existing reference template to identify the desired signal, we must develop our template from a training set.
We will discuss how we develop our set of templates for use in our matched-filter in section \ref{sec:training}.

@@

\begin{figure}[!ht]
   \centering
   \subfloat[\label{fig:multiple_comparison}]{\includegraphics[width=0.45\textwidth]{more_time_plots.png}}
   \quad
   \subfloat[\label{fig:timing_refs}]{\includegraphics[width=0.45\textwidth]{one_and_two_references.png}}
   \caption{Timing results with varying sampling frequencies, time-window lengths, and number of references used for the matched-filter on the MSP432 microcontroller processor. The time window used for (b) is 4.0 seconds with timings measured at sampling frequencies of $16, 32, 62, 64, 75, 100,$ and $128$ Hz.}
   \label{fig:timing}
\end{figure}
%
Because the accelerometer was not connected to the microcontroller at the time of writing this paper, randomized data was used for the purpose of analyzing the throughput performance on the device while varying the parameters which accurately models the processing load we would expect when we do have the accelerometer connected to the microcontroller.
Figure \ref{fig:multiple_comparison} shows the execution time with varying time-window lengths and sampling frequencies while maintaining two references used in the matched-filter process and operating at decimation $2$.
Because the FFT dominates the processing load, and was optimized for even-powered factors, one expects best throughput at radix 2 valued time-window lengths and sampling frequencies followed by other even-valued, non-radix 2, time-lengths and sampling frequencies. Both $62$Hz and $50$Hz have only one factor of $2$ hence the expectation that a sampling frequency of $62$Hz processing time is longer than that the $50$Hz.
Figure \ref{fig:timing_refs} shows the execution time over a $4.0$ second time window at decimation $8$ while varying the sampling frequency and number of references. Because the pre-processing execution time does not change with the number of references, it's observed that matched-filtering accounts for roughly $80$\% of the processing while pre-processing accounts for the other $20$\% when processing for a single reference.
%
\section{Conclusion}
We have demonstrated excellent activity classification capability of the matched-filter method using data obtained from a Kinetisense tri-axis accelerometer sensor.
We have shown that real-time processing can be achieved with timing results obtained from the MSP432 microcontroller.
Timing results have demonstrated that radix 2 time-window lengths and frequency values result in faster processing times as expected given that the FFT dominates the processing time.
We increased our throughput with the use of data-compression techniques, namely decimation and dimensionality reduction, which reduced our computational requirements.
The references used in the matched-filter are obtained from the training module where a training set is provided by the user through some specified activity and the best representative motion signature of the individual is extracted through the use of an instance-based learning algorithm.
